---
title: "STAT380 Final Project"
author: "Adam Abdel-Latif & Yuerong Shu"
date: "2026-04-27"
output: html_document
---

## Task 1

### Relevant Info

Prior to each online match, players in the game lobby are presented with two options for the battlefield of the upcoming game (`Map1` and`Map2`). The players have the option to vote, and the resulting vote is recorded in the `MapVote` column.
The winning map is listed in the `Choice` column. In the event of a tie vote, the map listed in `Map1` is chosen. (Games for which the player entered the lobby after the vote has taken place have no information in `Map1` and `Map2` but have the winning map presented in `Choice`.)

**Research Question:** Which maps are the most likely to win the map vote when they are an option?

**Strategy:** Find the rate for which each map is selected, both in total and when it is in the map pool. Ignore maps that are the "choice" when a player has entered the lobby after the vote has taken place. Find if there are any significant tie correlations between maps.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

p1 <- read.csv("CODGames_p1_380.csv", stringsAsFactors = FALSE)
p2 <- read.csv("CODGames_p2_380.csv", stringsAsFactors = FALSE)

cod <- bind_rows(p1, p2)
```

```{r}
# fix misspellings and inconsistent capitalisation in map name columns
fix_map_name <- function(x) {
  x <- trimws(x)
  case_when(
    x == "Drive-in"                           ~ "Drive-In",
    x %in% c("Apocolypse", "APocalypse")      ~ "Apocalypse",
    x %in% c("Collateral", "Collaterel Strike",
              "Collaterol Strike",
              "Collateral Striek")             ~ "Collateral Strike",
    x %in% c("Miami", "Miami Sstrike")        ~ "Miami Strike",
    x %in% c("Deisel", "Diesel")              ~ "Diesel",
    x == "Riad"                               ~ "Raid",
    x == "Ruah"                               ~ "Rush",
    x == "yamantau"                           ~ "Yamantau",
    x == "Deprogam"                           ~ "Deprogram",
    x == "Amrada Strike"                      ~ "Armada Strike",
    TRUE                                      ~ x
  )
}

cod <- cod %>%
  mutate(
    Choice = fix_map_name(Choice),
    Map1   = fix_map_name(Map1),
    Map2   = fix_map_name(Map2)
  )
```

```{r}
# remove games where player joined after the vote (Map1 and Map2 are blank)
voted_games <- cod %>%
  filter(Map1 != "" & Map2 != "")
```

The table below shows how often each map was selected as the winning map across all games in the dataset, including matches where the player joined after the vote had already taken place.

```{r}
# overall rate: how often each map was chosen out of all games
total_games <- nrow(cod)

overall_rate <- cod %>%
  count(Choice, name = "total_wins") %>%
  mutate(overall_rate = round(total_wins / total_games, 3)) %>%
  arrange(desc(overall_rate))

overall_rate
```

```{r}
# candidate rate: how often each map won when it was actually an option
candidates <- voted_games %>%
  pivot_longer(cols = c(Map1, Map2), names_to = "position", values_to = "map") %>%
  group_by(map) %>%
  summarise(
    times_candidate = n(),
    times_won = sum(map == Choice)
  ) %>%
  mutate(win_rate = round(times_won / times_candidate, 3))

# count wins that came from actual votes (strictly more votes, not ties)
vote_win_counts <- voted_games %>%
  filter(grepl("^[0-9]+ to [0-9]+$", MapVote)) %>%
  separate(MapVote, into = c("map1_votes", "map2_votes"), sep = " to ", convert = TRUE) %>%
  pivot_longer(cols = c(Map1, Map2), names_to = "position", values_to = "map") %>%
  mutate(
    my_votes = ifelse(position == "Map1", map1_votes, map2_votes),
    opp_votes = ifelse(position == "Map1", map2_votes, map1_votes),
    vote_win = (map == Choice) & (my_votes > opp_votes)
  ) %>%
  group_by(map) %>%
  summarise(vote_wins = sum(vote_win))

# join and compute both rates
candidates <- candidates %>%
  left_join(vote_win_counts, by = "map") %>%
  mutate(
    vote_wins = replace_na(vote_wins, 0),
    vote_win_rate = round(vote_wins / times_candidate, 3)
  ) %>%
  arrange(desc(win_rate))

candidates
```

```{r fig.align='center', fig.width=7, fig.height=5}
ggplot(candidates, aes(x = reorder(map, win_rate), y = win_rate)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Map Win Rate When Listed as a Candidate",
       x = "Map", y = "Win Rate") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_line(color = "gray80", linewidth = 0.5),
    panel.grid.minor.x = element_line(color = "gray90", linewidth = 0.3)
  )
```

```{r}
# tie analysis: find games where votes were equal (Map1 wins by default in a tie)
tie_games <- voted_games %>%
  filter(grepl("^[0-9]+ to [0-9]+$", MapVote)) %>%
  separate(MapVote, into = c("map1_votes", "map2_votes"), sep = " to ", convert = TRUE) %>%
  filter(map1_votes == map2_votes)

# which maps benefited most from tie wins
tie_summary <- tie_games %>%
  count(Map1, name = "tie_wins") %>%
  arrange(desc(tie_wins))

tie_summary
```

Based on the map win rate table, it would seem that Nuketown 84' is the most likely map to be chosen when presented as an option, followed by Crossroads Strike and Raid (all have above a 75% chance of being selected when listed as an option).
Although at first glance it appears that Nuketown 84' Halloween is the most likely to be chosen, it is only played once in our dataset and can be considered an insignificant outlier since it was likely a limited-time event map which had its own playlist and thus is the only option to play next.

When checking if ties could have an impact on map selection, an interesting thing I noticed is that the maps which have the most tie wins (they are map1 in the selection) are normally maps chosen less in the whole data set. Moscow, Cartel, Apocalypse, Zoo, and The Pines are all chosen less then 50% of the time when in the map selection pool and yet have the highest rates of ties. My hypothesis would be that ties most often occur when the options for the next map are generally disliked or viewed in indifference by the player population.

Based on the data, I think it is safe to say Nuketown '84 is the most popular map, even when accounting for ties and other factors. When it is an option, players chose it 82.1% of the time. 
## Task2

### Conversation with AI, prompts and results
Used generative AI: ChatGPT
Prompt1: I have uploaded 4 csv and here is the background of those datasets, get yourself familiar with those and we are going to do some data cleaning and EDA latter: [pasted the background introduction of all 4 datasets and attached csv files]
*ChatGPT replied nothing

Prompt2: Here is the instruction: [pasted the requirements of task1] Now help me write a complete solution of this question in R with annotation to major steps. You can solve things step by step and ask me what to do now while providing me options.
*ChatGPT replied:

Below is a complete R solution you can use.

Discussion / plan paragraph

To answer the research question, I will first combine the Player 1 and Player 2 datasets into one dataset because both contain the same type of match-level information. Since the research question focuses on map voting, I will use the variables Map1, Map2, Choice, and MapVote. Some rows have missing Map1 and Map2 because the player entered the lobby after voting was completed; those rows cannot be used to determine which maps were candidates, so they should be excluded from the vote-candidate analysis. I will also clean map names by trimming extra blanks and correcting misspellings using the official map names from CODMaps.csv as the reference.

For each match with two map options, I will count each listed map as one candidate appearance. A map will be counted as a winner when it matches the Choice column. I will also separate wins into two types: a direct vote win, where the map received more votes than the other option, and a tie-default win, where the vote was tied and the map won because it was listed as Map1. Finally, I will calculate the win proportion as number of wins / number of times candidate, then create a table and bar chart to show which maps were most likely to win when they appeared as an option.

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)

# Read in datasets
p1 <- read.csv("CODGames_p1_380.csv", stringsAsFactors = FALSE)
p2 <- read.csv("CODGames_p2_380.csv", stringsAsFactors = FALSE)
maps <- read.csv("CODMaps.csv", stringsAsFactors = FALSE)

# Combine player datasets
cod_games <- bind_rows(
  p1 %>% mutate(Player = "Player 1"),
  p2 %>% mutate(Player = "Player 2")
)

# Check official map names
maps$Name
```

```{r}
# Clean common map-name problems without editing the csv files
clean_map_name <- function(x) {
  x <- str_trim(x)

  x <- case_when(
    x %in% c("Collateral", "Collateral Striek", 
             "Collaterol Strike", "Collaterel Strike") ~ "Collateral Strike",
    x == "Deprogam" ~ "Deprogram",
    x == "Drive-in" ~ "Drive-In",
    x %in% c("Miami Stirke", "Miami Sstrike") ~ "Miami Strike",
    x == "Riad" ~ "Raid",
    x == "Ruah" ~ "Rush",
    x == "Amrada Strike" ~ "Armada Strike",
    x %in% c("APocalypse", "Apocolypse") ~ "Apocalypse",
    x == "Deisel" ~ "Diesel",
    x == "yamantau" ~ "Yamantau",
    x == "Nuketown '84 Halloween" ~ "Nuketown '84",
    TRUE ~ x
  )

  x
}

cod_games_clean <- cod_games %>%
  mutate(
    Map1 = clean_map_name(Map1),
    Map2 = clean_map_name(Map2),
    Choice = clean_map_name(Choice),
    MapVote = str_trim(MapVote),
    MapVote = str_replace(MapVote, " o ", " to ")
  )
```

```{r}
# Parse MapVote into votes for Map1 and Map2
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

```{r}
# Create one row per candidate map
map_candidates <- cod_votes %>%
  mutate(GameID = row_number()) %>%
  pivot_longer(
    cols = c(Map1, Map2),
    names_to = "MapPosition",
    values_to = "MapName"
  ) %>%
  mutate(
    Candidate = 1,
    Won = ifelse(MapName == Choice, 1, 0),

    CandidateVotes = ifelse(MapPosition == "Map1", Map1Votes, Map2Votes),
    OpponentVotes = ifelse(MapPosition == "Map1", Map2Votes, Map1Votes),

    VoteWin = ifelse(Won == 1 & CandidateVotes > OpponentVotes, 1, 0),
    TieDefaultWin = ifelse(Won == 1 & MapPosition == "Map1" &
                             CandidateVotes == OpponentVotes, 1, 0)
  )
```

```{r}
# Summarize by map
map_vote_summary <- map_candidates %>%
  group_by(MapName) %>%
  summarise(
    Times_Candidate = n(),
    Total_Wins = sum(Won),
    Vote_Wins = sum(VoteWin),
    Tie_Default_Wins = sum(TieDefaultWin),
    Win_Proportion = round(Total_Wins / Times_Candidate, 3),
    .groups = "drop"
  ) %>%
  arrange(desc(Win_Proportion))

map_vote_summary
```

```{r}
# Visualization
ggplot(map_vote_summary,
       aes(x = fct_reorder(MapName, Win_Proportion),
           y = Win_Proportion)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Probability of Winning Map Vote When Listed as a Candidate",
    x = "Map",
    y = "Win Proportion"
  ) +
  theme_minimal()
```
Answer to the research question

Based on the cleaned and combined data, the maps most likely to win the map vote when they were listed as an option were the maps with the highest Win_Proportion. In my calculation, the strongest maps were Nuketown ’84, Crossroads Strike, Raid, Diesel, and Standoff. These maps won a large proportion of the times they appeared as candidates. The table also separates wins earned by receiving more votes from wins caused by a tie where the map was chosen because it was listed as Map1, which is important because tie-default wins do not necessarily show stronger player preference.

Prompt3: I noticed that when I run the chunk # Parse MapVote into votes for Map1 and Map2, R send me this warning: Warning :Expected 2 pieces. Missing pieces filled with `NA` in 165 rows [3, 7, 9, 16, 18, 22, 23, 26, 42, 50, 56, 59, 61, 67, 71, 79, 85,
86, 99, 109, ...]. How would you deal with NA? Does this reveal some potential problem in your data cleansing process? 

*ChatGPT answered:
Good catch — this warning is important, and yes, it tells you something about your data cleaning.

What the warning means?
Expected 2 pieces. Missing pieces filled with NA
This happens because separate(..., sep = " to ") expected something like:
"3 to 5"
But in some rows, MapVote is:

missing (NA)
not in "X to Y" format
something like "Skipped", "—", "No vote", etc.
or badly formatted (extra spaces, typos)

So R cannot split into two numbers → produces NA.

Does this reveal a problem?
YES — but not necessarily a mistake

It reveals:

1. Data quality issues in MapVote

Some rows do not contain valid vote information

2. These rows should NOT be used for vote comparison

Because you cannot determine:

which map got more votes
whether it was a tie
Correct way to handle it (what your professor expects)

You should:

Step 1: Inspect problematic rows
```{r}
cod_games_clean %>%
  filter(!str_detect(MapVote, "to"))
```
This shows rows causing the issue.

Step 2: Keep only valid vote rows for vote-based analysis

Modify your code:
```{r}
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "to")) %>%   # <-- IMPORTANT FIX
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

Step 3: (Optional) keep track of excluded rows
```{r}
invalid_votes <- cod_games_clean %>%
  filter(!str_detect(MapVote, "to"))

nrow(invalid_votes)
```

Why this is correct statistically

Your research question requires:

map wins when it receives more votes

If you don't have vote counts, you cannot determine a vote-based win, so those rows must be excluded.

Very important nuance (for your write-up)

There are two types of missing cases:

Case A: Player joined after voting
Map1, Map2 missing → already filtered
Case B: Vote format missing/invalid
MapVote not usable → must also filter
What to write in your report (VERY important for points)

Some observations contained missing or improperly formatted vote information in the MapVote variable, which prevented extraction of vote counts. These rows were excluded from the vote-based analysis, as it was not possible to determine which map received more votes.

Bonus: More robust parsing (optional)

If you want to be extra safe:
```{r}
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "^[0-9]+ to [0-9]+$")) %>%
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

How to deal with NA?
Filter out rows where MapVote cannot be parsed
Does it reveal a problem?
Yes: inconsistent vote formatting / missing vote data
No: your cleaning is mostly correct — this is real-world messy data
My recommendation

Add the filter and mention it in your report → this is exactly the kind of reasoning instructors want to see.

If you want, next step we can:
👉 check how many rows you are losing and whether it biases results

Prompt4: Please do so and could you teach me how to keep a chunk with code in file without allowing it running when kniting document?

*ChatGPT replied: 
Part 1: Check how many rows are excluded and whether it may bias results

Use this chunk:
```{r}
# Total rows in combined dataset
n_total <- nrow(cod_games_clean)

# Rows with both candidate maps present
n_candidates <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  nrow()

# Rows with valid MapVote format
n_valid_vote <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "^[0-9]+ to [0-9]+$")) %>%
  nrow()

# Rows excluded because of invalid vote info
n_excluded <- n_candidates - n_valid_vote

# Percent excluded
pct_excluded <- round(100 * n_excluded / n_candidates, 2)

n_total
n_candidates
n_valid_vote
n_excluded
pct_excluded
```

Check if excluded rows are concentrated on certain maps
```{r}
excluded_rows <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(!str_detect(MapVote, "^[0-9]+ to [0-9]+$"))

table(excluded_rows$Choice)
```
If one map appears unusually often among excluded rows, mention possible bias.

Then it taught me way to prevent code running.

Prompt5: here are the result for some of the previous chunks: Are there anything we can improve? are we on the right track? [result of chunks pasted to analyze]

*ChatGPT replied:
```{r}
invalid_votes %>%
  count(MapVote, sort = TRUE)
```

This tells you whether they are:

blank
NA
weird formats like "Skipped"
"2to3"
"3 - 2"

This may allow recovery of some rows.

Improvement 2: Fix additional formatting before excluding

Sometimes rows fail because spacing inconsistent.

Use:
```{r}
cod_games_clean <- cod_games_clean %>%
  mutate(
    MapVote = str_trim(MapVote),
    MapVote = str_replace_all(MapVote, "\\s+", " "),
    MapVote = str_replace_all(MapVote, "-", " to "),
    MapVote = str_replace_all(MapVote, "to", " to ")
  )
```

Then rerun parsing.

This may reduce 165 exclusions.

Improvement 3: Check if exclusions bias certain maps

Your output shows:

Standoff = 14 excluded
Diesel = 10
Crossroads Strike = 9
Hijacked = 9
Armada Strike = 9

That suggests exclusions are not perfectly even.

Meaning some maps may be slightly undercounted.

This should be mentioned.

Write this in report

Approximately 19.7% of matches with candidate maps had unusable vote-count formatting and were excluded. Excluded matches were spread across many maps, although some maps such as Standoff and Diesel appeared more frequently. Therefore, final win probabilities should be interpreted with minor caution.

That sounds professional.

Improvement 4: VERY important — separate two research questions

Your assignment asks:

Which maps most likely to win when an option?

This can be answered TWO ways:

Method A (Best)

Use Choice

includes all 839 rows
directly tells winner
Method B

Use vote counts

only 674 rows
allows tie/default distinction
Best strategy:

Use 839 rows for candidate/winner proportions, then use 674 rows to study tie wins.

This is much stronger than using only 674.

### Discussion of prompt


### Comparison with result in task1

## Task 3 — Inference

### Relevant Information

The `GameType` variable records what mode was played each match. Hardpoint has teams fighting over a rotating location for points, while TDM is purely about getting kills. These modes play out completely differently, with different pacing, different time limits, and different ways to earn XP. My question here is whether the mode itself changes how much total XP you walk away with, even when you played equally well in both.

**Research Question:** How does the game type affect TotalXP after accounting for the player's Score?

---

### Data Cleaning

One thing I noticed is that the `GameType` column lists both normal and Hardcore (HC) versions of each mode separately (e.g., `”HC - TDM”` and `”TDM”`). Since there is no real distinction between them for this analysis, I stripped the `”HC - “` prefix to merge them. I also dropped any rows with missing values in `TotalXP`, `Score`, or `GameType`.

```{r message=FALSE, warning=FALSE}
library(stringr)

cod <- cod %>%
  mutate(GameType = str_trim(str_remove(GameType, "^HC\\s*-\\s*")))

cod_task3 <- cod %>%
  filter(!is.na(TotalXP), !is.na(Score), GameType != "")
```

After cleaning, the unique game types present in the data are:

```{r}
unique(cod_task3$GameType)
```

---

### Exploratory Data Analysis

Before fitting anything, it helps to look at how `TotalXP` and `Score` are distributed across game types to get a sense of whether the differences are large enough to matter.

#### Summary Statistics

The table below breaks down mean, median, and standard deviation of both `TotalXP` and `Score` for each game type.

```{r message=FALSE}
task3_stats <- cod_task3 %>%
  group_by(GameType) %>%
  summarise(
    n            = n(),
    Mean_XP      = round(mean(TotalXP), 0),
    Median_XP    = round(median(TotalXP), 0),
    SD_XP        = round(sd(TotalXP), 0),
    Mean_Score   = round(mean(Score), 0),
    Median_Score = round(median(Score), 0),
    SD_Score     = round(sd(Score), 0)
  ) %>%
  arrange(desc(Mean_XP))

task3_stats
```

Looking at the table, it would seem that Hardpoint and Domination both average noticeably more XP than TDM and Kill Confirmed, which makes sense since those modes involve more sustained team activity and likely run longer. An interesting thing I noticed is that Score also varies across modes, with Domination scores clustering much more tightly than the others. This matters because if I just compared raw XP across modes without accounting for Score, the results would be mixing up the effect of game type with how well the player performed, which is why Score needs to be in the model. It is also worth noting that Domination only has 14 observations in this dataset, so its estimates should be interpreted with some caution.

---

#### Distribution of TotalXP by Game Type

```{r fig.align='center', fig.width=7, fig.height=4}
ggplot(cod_task3, aes(x = reorder(GameType, TotalXP, median), y = TotalXP)) +
  geom_boxplot(fill = "steelblue", color = "gray30", outlier.alpha = 0.3) +
  coord_flip() +
  labs(
    title = "Distribution of Total XP by Game Type",
    x     = "Game Type",
    y     = "Total XP"
  ) +
  theme_minimal(base_size = 12)
```

The boxplots make the differences fairly clear. Domination and Hardpoint have noticeably higher median XP, while Kill Confirmed sits at the bottom. There is also a lot of spread within each mode though, which suggests game type alone does not tell the whole story, and how well you personally played clearly matters too.

---

#### Distribution of Score by Game Type

```{r fig.align='center', fig.width=7, fig.height=4}
ggplot(cod_task3, aes(x = reorder(GameType, Score, median), y = Score)) +
  geom_boxplot(fill = "coral", color = "gray30", outlier.alpha = 0.3) +
  coord_flip() +
  labs(
    title = "Distribution of Score by Game Type",
    x     = "Game Type",
    y     = "Score"
  ) +
  theme_minimal(base_size = 12)
```

This confirms what the summary table hinted at. Score is not spread evenly across game types, with Domination in particular having a much narrower score range than Hardpoint or TDM. If I only looked at raw XP differences across modes without accounting for Score, the results would be misleading since the modes naturally produce different score distributions to begin with.

---

#### TotalXP vs. Score by Game Type

```{r fig.align='center', fig.width=8, fig.height=5, message=FALSE, warning=FALSE}
ggplot(cod_task3, aes(x = Score, y = TotalXP, color = GameType)) +
  geom_point(alpha = 0.2, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.9) +
  labs(
    title  = "Total XP vs. Score by Game Type",
    x      = "Score",
    y      = "Total XP",
    color  = "Game Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right")
```

This is probably the most useful plot in the EDA. There is a clear positive relationship between Score and TotalXP in every game type, which makes sense since better performance means more XP. An interesting thing I noticed is that the slopes look quite different across modes. Hardpoint has a much steeper trend line than TDM or Kill Confirmed, meaning each extra point of Score earns more XP in Hardpoint than it would in other modes. My hypothesis is that this points to a possible interaction between Score and GameType, though for this analysis I will fit the simpler additive model since it is more interpretable.

---

### Model: Multiple Linear Regression

I fit a multiple linear regression with `TotalXP` as the response, `Score` as a continuous predictor, and `GameType` as a categorical one. Putting both in the same model lets Score be held constant, which isolates what game type is actually contributing to XP on its own.

```{r}
model_task3 <- lm(TotalXP ~ Score + GameType, data = cod_task3)
summary(model_task3)
```

The `anova()` output runs a partial F-test for `GameType`, which tells us whether adding game type to the model actually helps explain TotalXP beyond what Score already covers.

```{r}
anova(model_task3)
```

---

### Model Assumptions

Before reading too much into the coefficients, it's worth checking the residual plots to see if anything looks off.

```{r fig.align='center', fig.width=9, fig.height=4}
par(mfrow = c(1, 2))

plot(model_task3, which = 1, main = "Residuals vs Fitted")
plot(model_task3, which = 2, main = "Normal Q-Q")

par(mfrow = c(1, 1))
```

The residuals vs. fitted plot has a bit of a fan shape, with the spread getting wider as fitted values increase. This suggests some heteroscedasticity, meaning the variance is not perfectly constant across the range of predictions. It is not too surprising given that high-XP games tend to be harder to predict. It does not break the model, but it is worth being a little cautious about taking the p-values too literally. The QQ-plot looks fairly good through the middle of the data but drifts a bit at the upper tail, which is consistent with a handful of unusually high-XP outlier games. With the sample size available here, inference should still hold reasonably well.

---

### Answer to the Research Question

Based on the model output, it would seem that game type does have a statistically significant effect on TotalXP even after accounting for Score, with the partial F-test returning F = 14.05 and p = 6.23e-09. Score is doing most of the heavy lifting though, which makes sense. The Score coefficient is `r round(coef(model_task3)["Score"], 2)`, meaning each additional point of Score adds roughly that many XP regardless of mode. An interesting thing I noticed is that even though GameType clears the significance bar as a group, none of the individual game type levels differ significantly from Domination on their own. Kill Confirmed and TDM both trend negative, which lines up with what the EDA showed since kill-based modes tend to produce less XP than objective ones, but neither individually reaches the 0.05 threshold. I think it is safe to say that game type does matter for TotalXP independently of performance, just that the effect is spread across modes rather than being driven by any one in particular.


## Task 4 — Prediction

**Research Question:** Which classification method best predicts match outcome (Win/Loss) using Score, Eliminations, Deaths, and GameType as predictors?

### Data Wrangling

The `Result` column does not just say "Win" or "Loss", it actually stores the match score like `"100-97"`. To get a usable outcome variable, I pulled out the two numbers and compared them to determine who won. I also filtered to only full matches since partial games do not have a complete result, and dropped ties since they do not count as a win or a loss.

```{r message=FALSE, warning=FALSE}
library(randomForest)
library(stringr)

cod_t4 <- cod %>%
  filter(FullPartial == "Full", Result != "", !is.na(Score),
         !is.na(Eliminations), !is.na(Deaths)) %>%
  mutate(
    team_score = as.numeric(str_extract(Result, "^[0-9]+")),
    opp_score  = as.numeric(str_extract(Result, "[0-9]+$")),
    Outcome = case_when(
      team_score > opp_score ~ "Win",
      team_score < opp_score ~ "Loss",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Outcome)) %>%
  mutate(
    Outcome  = factor(Outcome),
    GameType = factor(GameType)
  ) %>%
  select(Outcome, Score, Eliminations, Deaths, GameType)

table(cod_t4$Outcome)
```

It is worth checking the class balance before fitting anything. If one outcome is way more common than the other, a model could just predict that class every time and still look accurate on paper, which would not actually be useful. The split here is roughly 403 wins to 318 losses, which is a little uneven but not bad enough to cause major problems for the model.

---

### Train/Test Split

I split the data 80/20 into training and test sets. The seed ensures the split is the same every time the code is run, and it also means the comparison with logistic regression will use the exact same test observations.

```{r}
set.seed(380)
train_idx <- sample(nrow(cod_t4), size = 0.8 * nrow(cod_t4))
train_t4  <- cod_t4[train_idx, ]
test_t4   <- cod_t4[-train_idx, ]
```

---

### Random Forest

Random Forest builds a large number of decision trees, each trained on a different random sample of the training data. At every split, only a random subset of predictors is considered rather than all of them. That second part is what makes it work well. If every tree saw the same predictors at every split, they would all end up looking similar and making the same errors. By forcing each tree to work with different information, you end up with trees that each make different mistakes, and averaging across them cancels a lot of those errors out. Predicting a new observation is just a vote, with whichever class gets the most votes across all 500 trees winning. It tends to be much more accurate than a single decision tree and does not need much preprocessing.

#### Choosing mtry

The main tuning parameter for Random Forest is `mtry`, which controls how many predictors are randomly considered at each split. The default for classification is the square root of the number of predictors, so with 4 predictors that is 2. I used `tuneRF` to check whether a different value performs better based on the out-of-bag error.

```{r fig.align='center', fig.width=6, fig.height=4}
set.seed(380)
tuneRF(
  x          = train_t4[, -1],
  y          = train_t4$Outcome,
  stepFactor = 1.5,
  improve    = 0.01,
  plot       = TRUE,
  trace      = FALSE
)
```

Based on the plot, `mtry = 2` gives the lowest OOB error, so that is what I used in the final model.

#### Model

```{r}
set.seed(380)
rf_model <- randomForest(
  Outcome ~ Score + Eliminations + Deaths + GameType,
  data       = train_t4,
  mtry       = 2,
  ntree      = 500,
  importance = TRUE
)

rf_model
```

The OOB error shown above is essentially the model evaluating itself on observations it was not trained on, so it gives a reasonable early estimate of how well it will perform on the actual test set.

---

### Test Set Results

```{r}
rf_preds <- predict(rf_model, newdata = test_t4)

table(Predicted = rf_preds, Actual = test_t4$Outcome)

rf_accuracy <- mean(rf_preds == test_t4$Outcome)
cat("Random Forest Accuracy:", round(rf_accuracy, 4))
```

62% accuracy is better than guessing, though it is not a particularly strong result. Looking at the confusion matrix, the model correctly predicted 42 losses and 48 wins, but misclassified 29 actual wins as losses and 26 actual losses as wins. It is making errors in both directions fairly evenly, so it is not simply defaulting to predicting one class.

---

### Variable Importance

One useful thing about Random Forest is that it can tell us which predictors actually mattered. The left panel shows how much accuracy drops when a variable is shuffled randomly, with a larger drop indicating the model relied on it more heavily. The right panel measures how much each variable reduced impurity across all the splits in all 500 trees.

```{r fig.align='center', fig.width=7, fig.height=4}
varImpPlot(rf_model, main = "Variable Importance — Random Forest")
```

An interesting thing I noticed is that Deaths ranked as the most important predictor by MeanDecreaseAccuracy, which makes sense since dying frequently puts your team at a disadvantage regardless of what mode you are playing. Score actually ranked last by that same metric, even though it ranked highest in MeanDecreaseGini. The two metrics disagreeing on Score is worth noting. My hypothesis is that Score helps the trees split the data cleanly, but removing it entirely does not hurt overall accuracy as much as removing Deaths would. GameType ranked last by MeanDecreaseGini, which would suggest that win or loss comes down more to individual performance than to which game mode was being played.

---

## Task4 Part 2
Research Question: 
"Which classification method best predicts match outcome (Win/Loss) using Score, Eliminations, Deaths, and GameType as predictors?"

We decided to choose logistic regression out of all learned classification method.

### Logistic Regression:

Chunk1: This chunk filters the dataset to include only complete matches and removes missing values in key variables. It then extracts the player’s score and opponent’s score from the Result column. The outcome variable is created by comparing these scores: if the player’s score is higher, the outcome is a win; otherwise, it is a loss.

```{r}
library(dplyr)
library(stringr)
library(pROC)

cod_t4 <- cod %>%
  filter(
    FullPartial == "Full",
    Result != "",
    !is.na(Score),
    !is.na(Eliminations),
    !is.na(Deaths)
  ) %>%
  mutate(
    team_score = as.numeric(str_extract(Result, "^[0-9]+")),
    opp_score  = as.numeric(str_extract(Result, "[0-9]+$")),
    Outcome = case_when(
      team_score > opp_score ~ "Win",
      team_score < opp_score ~ "Loss",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Outcome)) %>%
  mutate(
    Outcome  = factor(Outcome, levels = c("Loss", "Win")),
    GameType = factor(GameType)
  ) %>%
  select(Outcome, Score, Eliminations, Deaths, GameType)

table(cod_t4$Outcome)
```
Chunk 2: Create training and validation sets

This chunk splits the data into a training set (80%) and a validation set (20%). The training set is used to fit the model, while the validation set is used to evaluate predictive performance on unseen data.
```{r}
set.seed(123)

train_ind <- sample(
  1:nrow(cod_t4),
  floor(0.8 * nrow(cod_t4))
)

Train <- cod_t4[train_ind, ]
Validation <- cod_t4[-train_ind, ]
```

Chunk 3: Fit the logistic regression model

This chunk fits a logistic regression model using the training data. The model estimates the probability of winning based on Score, Eliminations, Deaths, and GameType， the variables we choose as important according to game experience.

```{r}
log_mod <- glm(
  Outcome ~ Score + Eliminations + Deaths + GameType,
  data = Train,
  family = binomial
)

summary(log_mod)
```
From the model summary:

Eliminations: Higher eliminations are associated with a higher probability of winning.
Deaths: More deaths significantly decrease the probability of winning.
Score: does not appear to add much predictive value once eliminations and deaths are included.
GameType: does not have a meaningful effect on match outcome in this model.
Even though Score seems intuitively important, it is not statistically significant in the model. This is likely because:

We can see that Score is highly correlated with eliminations and deaths and the model already captures performance through those variables

Chunk 4: Predict win probabilities

This chunk uses the fitted logistic regression model to predict the probability of a win for each observation in the validation set. The type = "response" option returns predicted probabilities instead of log-odds.
```{r}
log_prob <- predict(
  log_mod,
  newdata = Validation,
  type = "response"
)
```

Chunk 5: Convert probabilities into predicted classes

This chunk converts predicted probabilities into predicted match outcomes. I used a cutoff of 0.5: if the predicted probability of winning is at least 0.5, the prediction is "Win"; otherwise, it is "Loss".

```{r}
log_pred <- ifelse(log_prob >= 0.5, "Win", "Loss")

log_pred <- factor(
  log_pred,
  levels = levels(Validation$Outcome)
)
```

Chunk 6: Confusion matrix and accuracy

This chunk evaluates the model using a confusion matrix and accuracy. The confusion matrix shows how many wins and losses were classified correctly or incorrectly. Accuracy gives the proportion of validation observations that were predicted correctly.

```{r}
log_conf <- table(
  Predicted = log_pred,
  Actual = Validation$Outcome
)

log_accuracy <- mean(log_pred == Validation$Outcome)

log_conf
log_accuracy
```
The logistic regression model shows moderate predictive performance. The validation accuracy is approximately 0.669, meaning that about 66.9% of match outcomes are correctly classified. This indicates that the model performs better than random guessing(50%), but there is still room for improvement.

Chunk 7: ROC curve and AUC

This chunk calculates the ROC curve and AUC. AUC measures how well the model separates wins from losses across possible probability cutoffs. A larger AUC means the model is better at distinguishing wins from losses.
```{r}
log_roc <- roc(
  Validation$Outcome,
  log_prob,
  levels = c("Loss", "Win")
)

log_auc <- auc(log_roc)

log_auc

plot(log_roc)
```
The AUC value is 0.7104, which suggests that the model has a reasonable ability to distinguish between wins and losses. In general, an AUC between 0.7 and 0.8 is considered acceptable, so this model demonstrates fair classification performance, though it is not highly accurate.

In a conclusion, the logistic regression model provides a reasonable but not strong prediction of match outcomes. While the model performs better than chance and demonstrates acceptable discrimination (AUC ≈ 0.71), its accuracy suggests that additional predictors or more complex models may improve performance. Among the predictors, eliminations and deaths are the most important factors influencing match outcome, while score and game type do not contribute significantly after accounting for these variables.
