---
title: "STAT380 Final Project"
author: "Adam Abdel-Latif & Yuerong Shu"
date: "2026-04-27"
output: html_document
---

## Task1

#Relevant Info
Prior to each online match, players inthe game lobby are presented with two options for the battlefield of the upcoming game (`Map1` and`Map2`). The players have the option to vote, and the resulting vote is recorded in the `MapVote` column.
The winning map is listed in the `Choice` column. In the event of a tie vote, the map listed in `Map1` is chosen. (Games for which the player entered the lobby after the vote has taken place have no information in `Map1` and `Map2` but have the winning map presented in `Choice`.)
Research Question: Which maps are the most likely to win the map vote when they are an option?

Strategy: Find the rate for which each map is selected, both in total and when it is in the map pool. Ignore maps that are the "choice" when a player has entered the lobby after the vote has taken place. Find if there are any significant tie correlations between maps. 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

p1 <- read.csv("CODGames_p1_380.csv", stringsAsFactors = FALSE)
p2 <- read.csv("CODGames_p2_380.csv", stringsAsFactors = FALSE)

cod <- bind_rows(p1, p2)
```

```{r}
# remove games where player joined after the vote (Map1 and Map2 are blank)
voted_games <- cod %>%
  filter(Map1 != "" & Map2 != "")
```

```{r}
# overall rate: how often each map was chosen out of all games
total_games <- nrow(cod)

overall_rate <- cod %>%
  count(Choice, name = "total_wins") %>%
  mutate(overall_rate = round(total_wins / total_games, 3)) %>%
  arrange(desc(overall_rate))

overall_rate
```

```{r}
# candidate rate: how often each map won when it was actually an option
candidates <- voted_games %>%
  pivot_longer(cols = c(Map1, Map2), names_to = "position", values_to = "map") %>%
  group_by(map) %>%
  summarise(
    times_candidate = n(),
    times_won = sum(map == Choice)
  ) %>%
  mutate(win_rate = round(times_won / times_candidate, 3))

# count wins that came from actual votes (strictly more votes, not ties)
vote_win_counts <- voted_games %>%
  filter(grepl("^[0-9]+ to [0-9]+$", MapVote)) %>%
  separate(MapVote, into = c("map1_votes", "map2_votes"), sep = " to ", convert = TRUE) %>%
  pivot_longer(cols = c(Map1, Map2), names_to = "position", values_to = "map") %>%
  mutate(
    my_votes = ifelse(position == "Map1", map1_votes, map2_votes),
    opp_votes = ifelse(position == "Map1", map2_votes, map1_votes),
    vote_win = (map == Choice) & (my_votes > opp_votes)
  ) %>%
  group_by(map) %>%
  summarise(vote_wins = sum(vote_win))

# join and compute both rates
candidates <- candidates %>%
  left_join(vote_win_counts, by = "map") %>%
  mutate(
    vote_wins = replace_na(vote_wins, 0),
    vote_win_rate = round(vote_wins / times_candidate, 3)
  ) %>%
  arrange(desc(win_rate))

candidates
```

```{r}
ggplot(candidates, aes(x = reorder(map, win_rate), y = win_rate)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Map Win Rate When Listed as a Candidate",
       x = "Map", y = "Win Rate") +
  theme_minimal()
```

```{r}
# tie analysis: find games where votes were equal (Map1 wins by default in a tie)
tie_games <- voted_games %>%
  filter(grepl("^[0-9]+ to [0-9]+$", MapVote)) %>%
  separate(MapVote, into = c("map1_votes", "map2_votes"), sep = " to ", convert = TRUE) %>%
  filter(map1_votes == map2_votes)

# which maps benefited most from tie wins
tie_summary <- tie_games %>%
  count(Map1, name = "tie_wins") %>%
  arrange(desc(tie_wins))

tie_summary
```

Based on the map win rate table, it would seem that Nuketown 84' is the most likely map to be chosen when presented as an option, followed by Crossroads Strike and Raid (all have above a 75% chance of being selected when listed as an option). 
Although at first glance it appears that Nuketown 84' Halloween is the most likely to be chosen, it is only played once in our dataset and can be considered an insignificant outlier since it was likely a limited-time event map which had its own playlist and thus is the only option to play next. 

When checking if ties could have an impact on map selection, an interesting thing I noticed is that the maps which have the most tie wins (they are map1 in the selection) are normally maps chosen less in the whole data set. Moscow, Cartel, Apocalypse, Zoo, and The Pines are all chosen less then 50% of the time when in the map selection pool and yet have the highest rates of ties. My hypothesis would be that ties most often occur when the options for the next map are generally disliked or viewed in indifference by the player population. 

This all being said, it is worth noting that the map Raid only had one tie in the entire dataset. Along with its regular win rate of 77.6%, it could be deduced that this is the most popular map when presented as an option and is the "best" map. 

## Task2

### Conversation with AI, prompts and results
Used generative AI: ChatGPT
Prompt1: I have uploaded 4 csv and here is the background of those datasets, get yourself familiar with those and we are going to do some data cleaning and EDA latter: [pasted the background introduction of all 4 datasets and attached csv files]
*ChatGPT replied nothing

Prompt2: Here is the instruction: [pasted the requirements of task1] Now help me write a complete solution of this question in R with annotation to major steps. You can solve things step by step and ask me what to do now while providing me options.
*ChatGPT replied:

Below is a complete R solution you can use.

Discussion / plan paragraph

To answer the research question, I will first combine the Player 1 and Player 2 datasets into one dataset because both contain the same type of match-level information. Since the research question focuses on map voting, I will use the variables Map1, Map2, Choice, and MapVote. Some rows have missing Map1 and Map2 because the player entered the lobby after voting was completed; those rows cannot be used to determine which maps were candidates, so they should be excluded from the vote-candidate analysis. I will also clean map names by trimming extra blanks and correcting misspellings using the official map names from CODMaps.csv as the reference.

For each match with two map options, I will count each listed map as one candidate appearance. A map will be counted as a winner when it matches the Choice column. I will also separate wins into two types: a direct vote win, where the map received more votes than the other option, and a tie-default win, where the vote was tied and the map won because it was listed as Map1. Finally, I will calculate the win proportion as number of wins / number of times candidate, then create a table and bar chart to show which maps were most likely to win when they appeared as an option.

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)

# Read in datasets
p1 <- read.csv("C:/Users/alice/Downloads/CODGames_p1_380.csv", stringsAsFactors = FALSE)
p2 <- read.csv("C:/Users/alice/Downloads/CODGames_p2_380.csv", stringsAsFactors = FALSE)
maps <- read.csv("C:/Users/alice/Downloads/CODMaps.csv", stringsAsFactors = FALSE)

# Combine player datasets
cod_games <- bind_rows(
  p1 %>% mutate(Player = "Player 1"),
  p2 %>% mutate(Player = "Player 2")
)

# Check official map names
maps$Name
```

```{r}
# Clean common map-name problems without editing the csv files
clean_map_name <- function(x) {
  x <- str_trim(x)

  x <- case_when(
    x %in% c("Collateral", "Collateral Striek", 
             "Collaterol Strike", "Collaterel Strike") ~ "Collateral Strike",
    x == "Deprogam" ~ "Deprogram",
    x == "Drive-in" ~ "Drive-In",
    x %in% c("Miami Stirke", "Miami Sstrike") ~ "Miami Strike",
    x == "Riad" ~ "Raid",
    x == "Ruah" ~ "Rush",
    x == "Amrada Strike" ~ "Armada Strike",
    x %in% c("APocalypse", "Apocolypse") ~ "Apocalypse",
    x == "Deisel" ~ "Diesel",
    x == "yamantau" ~ "Yamantau",
    x == "Nuketown '84 Halloween" ~ "Nuketown '84",
    TRUE ~ x
  )

  x
}

cod_games_clean <- cod_games %>%
  mutate(
    Map1 = clean_map_name(Map1),
    Map2 = clean_map_name(Map2),
    Choice = clean_map_name(Choice),
    MapVote = str_trim(MapVote),
    MapVote = str_replace(MapVote, " o ", " to ")
  )
```

```{r}
# Parse MapVote into votes for Map1 and Map2
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

```{r}
# Create one row per candidate map
map_candidates <- cod_votes %>%
  mutate(GameID = row_number()) %>%
  pivot_longer(
    cols = c(Map1, Map2),
    names_to = "MapPosition",
    values_to = "MapName"
  ) %>%
  mutate(
    Candidate = 1,
    Won = ifelse(MapName == Choice, 1, 0),

    CandidateVotes = ifelse(MapPosition == "Map1", Map1Votes, Map2Votes),
    OpponentVotes = ifelse(MapPosition == "Map1", Map2Votes, Map1Votes),

    VoteWin = ifelse(Won == 1 & CandidateVotes > OpponentVotes, 1, 0),
    TieDefaultWin = ifelse(Won == 1 & MapPosition == "Map1" &
                             CandidateVotes == OpponentVotes, 1, 0)
  )
```

```{r}
# Summarize by map
map_vote_summary <- map_candidates %>%
  group_by(MapName) %>%
  summarise(
    Times_Candidate = n(),
    Total_Wins = sum(Won),
    Vote_Wins = sum(VoteWin),
    Tie_Default_Wins = sum(TieDefaultWin),
    Win_Proportion = round(Total_Wins / Times_Candidate, 3),
    .groups = "drop"
  ) %>%
  arrange(desc(Win_Proportion))

map_vote_summary
```

```{r}
# Visualization
ggplot(map_vote_summary,
       aes(x = fct_reorder(MapName, Win_Proportion),
           y = Win_Proportion)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Probability of Winning Map Vote When Listed as a Candidate",
    x = "Map",
    y = "Win Proportion"
  ) +
  theme_minimal()
```
Answer to the research question

Based on the cleaned and combined data, the maps most likely to win the map vote when they were listed as an option were the maps with the highest Win_Proportion. In my calculation, the strongest maps were Nuketown ’84, Crossroads Strike, Raid, Diesel, and Standoff. These maps won a large proportion of the times they appeared as candidates. The table also separates wins earned by receiving more votes from wins caused by a tie where the map was chosen because it was listed as Map1, which is important because tie-default wins do not necessarily show stronger player preference.

Prompt3: I noticed that when I run the chunk # Parse MapVote into votes for Map1 and Map2, R send me this warning: Warning :Expected 2 pieces. Missing pieces filled with `NA` in 165 rows [3, 7, 9, 16, 18, 22, 23, 26, 42, 50, 56, 59, 61, 67, 71, 79, 85,
86, 99, 109, ...]. How would you deal with NA? Does this reveal some potential problem in your data cleansing process? 

*ChatGPT answered:
Good catch — this warning is important, and yes, it tells you something about your data cleaning.

What the warning means?
Expected 2 pieces. Missing pieces filled with NA
This happens because separate(..., sep = " to ") expected something like:
"3 to 5"
But in some rows, MapVote is:

missing (NA)
not in "X to Y" format
something like "Skipped", "—", "No vote", etc.
or badly formatted (extra spaces, typos)

So R cannot split into two numbers → produces NA.

Does this reveal a problem?
YES — but not necessarily a mistake

It reveals:

1. Data quality issues in MapVote

Some rows do not contain valid vote information

2. These rows should NOT be used for vote comparison

Because you cannot determine:

which map got more votes
whether it was a tie
Correct way to handle it (what your professor expects)

You should:

Step 1: Inspect problematic rows
```{r}
cod_games_clean %>%
  filter(!str_detect(MapVote, "to"))
```
This shows rows causing the issue.

Step 2: Keep only valid vote rows for vote-based analysis

Modify your code:
```{r}
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "to")) %>%   # <-- IMPORTANT FIX
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

Step 3: (Optional) keep track of excluded rows
```{r}
invalid_votes <- cod_games_clean %>%
  filter(!str_detect(MapVote, "to"))

nrow(invalid_votes)
```

Why this is correct statistically

Your research question requires:

map wins when it receives more votes

If you don't have vote counts, you cannot determine a vote-based win, so those rows must be excluded.

Very important nuance (for your write-up)

There are two types of missing cases:

Case A: Player joined after voting
Map1, Map2 missing → already filtered
Case B: Vote format missing/invalid
MapVote not usable → must also filter
What to write in your report (VERY important for points)

Some observations contained missing or improperly formatted vote information in the MapVote variable, which prevented extraction of vote counts. These rows were excluded from the vote-based analysis, as it was not possible to determine which map received more votes.

Bonus: More robust parsing (optional)

If you want to be extra safe:
```{r}
cod_votes <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "^[0-9]+ to [0-9]+$")) %>%
  separate(
    MapVote,
    into = c("Map1Votes", "Map2Votes"),
    sep = " to ",
    convert = TRUE
  )
```

How to deal with NA?
Filter out rows where MapVote cannot be parsed
Does it reveal a problem?
Yes: inconsistent vote formatting / missing vote data
No: your cleaning is mostly correct — this is real-world messy data
My recommendation

Add the filter and mention it in your report → this is exactly the kind of reasoning instructors want to see.

If you want, next step we can:
👉 check how many rows you are losing and whether it biases results

Prompt4: Please do so and could you teach me how to keep a chunk with code in file without allowing it running when kniting document?

*ChatGPT replied: 
Part 1: Check how many rows are excluded and whether it may bias results

Use this chunk:
```{r}
# Total rows in combined dataset
n_total <- nrow(cod_games_clean)

# Rows with both candidate maps present
n_candidates <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  nrow()

# Rows with valid MapVote format
n_valid_vote <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(str_detect(MapVote, "^[0-9]+ to [0-9]+$")) %>%
  nrow()

# Rows excluded because of invalid vote info
n_excluded <- n_candidates - n_valid_vote

# Percent excluded
pct_excluded <- round(100 * n_excluded / n_candidates, 2)

n_total
n_candidates
n_valid_vote
n_excluded
pct_excluded
```

Check if excluded rows are concentrated on certain maps
```{r}
excluded_rows <- cod_games_clean %>%
  filter(!is.na(Map1), !is.na(Map2)) %>%
  filter(!str_detect(MapVote, "^[0-9]+ to [0-9]+$"))

table(excluded_rows$Choice)
```
If one map appears unusually often among excluded rows, mention possible bias.

Then it taught me way to prevent code running.

Prompt5: here are the result for some of the previous chunks: Are there anything we can improve? are we on the right track? [result of chunks pasted to analyze]

*ChatGPT replied:
```{r}
invalid_votes %>%
  count(MapVote, sort = TRUE)
```

This tells you whether they are:

blank
NA
weird formats like "Skipped"
"2to3"
"3 - 2"

This may allow recovery of some rows.

Improvement 2: Fix additional formatting before excluding

Sometimes rows fail because spacing inconsistent.

Use:
```{r}
cod_games_clean <- cod_games_clean %>%
  mutate(
    MapVote = str_trim(MapVote),
    MapVote = str_replace_all(MapVote, "\\s+", " "),
    MapVote = str_replace_all(MapVote, "-", " to "),
    MapVote = str_replace_all(MapVote, "to", " to ")
  )
```

Then rerun parsing.

This may reduce 165 exclusions.

Improvement 3: Check if exclusions bias certain maps

Your output shows:

Standoff = 14 excluded
Diesel = 10
Crossroads Strike = 9
Hijacked = 9
Armada Strike = 9

That suggests exclusions are not perfectly even.

Meaning some maps may be slightly undercounted.

This should be mentioned.

Write this in report

Approximately 19.7% of matches with candidate maps had unusable vote-count formatting and were excluded. Excluded matches were spread across many maps, although some maps such as Standoff and Diesel appeared more frequently. Therefore, final win probabilities should be interpreted with minor caution.

That sounds professional.

Improvement 4: VERY important — separate two research questions

Your assignment asks:

Which maps most likely to win when an option?

This can be answered TWO ways:

Method A (Best)

Use Choice

includes all 839 rows
directly tells winner
Method B

Use vote counts

only 674 rows
allows tie/default distinction
Best strategy:

Use 839 rows for candidate/winner proportions, then use 674 rows to study tie wins.

This is much stronger than using only 674.

### Discussion of prompt


### Comparison with result in task1

## Task3



## Task4

